---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cobrar

<!-- badges: start -->
[![R-CMD-check](https://github.com/Waschina/cobrar/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Waschina/cobrar/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The R package *cobrar* provides structures and functions for constraint-based metabolic network analysis, e.g. the prediction of metabolic fluxes using fluxes using Flux Balance Analysis (FBA). *cobrar* is inspired by the former by the former CRAN R package *sybil*[(1)](#R1).

## Installation

Please note that *cobrar* requires the two system libraries *libSBML* and *glpk*. The following installation instructions for different operating systems install the dependencies first, then *cobrar*. If you already have *libSBML* and *glpk* installed, you can skip to the last part of the instructions.

*cobrar* is under development and the installation instructions described will install the latest development version.

#### Ubuntu/Debian/Mint

Install *libSBML* and *glpk*:
```sh
sudo apt install libsbml-dev libglpk-dev
```

Install cobrar (in *R*):
``` r
# install.packages("remotes")
remotes::install_github("Waschina/cobrar")
```

#### Centos/Fedora/RHEL

Install *libSBML* and *glpk*:
```sh
sudo yum install libsbml-devel glpk-devel
```

Install cobrar (in *R*):
``` r
# install.packages("remotes")
remotes::install_github("Waschina/cobrar")
```
#### MacOS

*libSBML* can be installed from source:
```sh
curl -LO https://github.com/sbmlteam/libsbml/archive/refs/tags/v5.20.2.tar.gz
tar -zxf v5.20.2.tar.gz
cd libsbml-5.20.2
mkdir build
cd build
cmake -DENABLE_FBC=ON -DENABLE_GROUPS=ON ..
make
sudo make install 
```

Install *glpk* either homebrew or MacPorts:
``` sh
# Homebrew
brew install glpk

# OR: MacPorts
sudo port install glpk
```

Add the the path to libSBML to the library paths that are checked by R.
``` sh
export R_LD_LIBRARY_PATH="`R RHOME`/lib:/usr/local/lib"
```

Install cobrar (in *R*):
``` r
# install.packages("remotes")
remotes::install_github("Waschina/cobrar")
``` 

Using *cobrar* in R:
``` r
dyn.load("/usr/local/lib/libsbml.5.20.2.dylib")
library(cobrar)
```

#### Windows
TODO

#### Conda
TODO

## Usage

The full documentation including illustrative examples is available [here](https://waschina.github.io/cobrar/).

The following (very brief) example illustrates how to use `cobrar` to perform flux balance analysis for a core metabolism model of *Escherichia coli*. In a first simulation, fluxes and growth are predicted for the default constraints of the model, which represents a minimal medium with glucose as the sole carbon source and the presence of oxygen. The second simulation performs the same flux balance analysis but without oxygen to simulate an anoxic growth environment.


```{r example}
# First, load the cobrar package and the metabolic model:
library(cobrar)
fpath <- system.file("extdata", "e_coli_core.xml", package="cobrar")
mod <- readSBMLmod(fpath)

# Next, perform simulation 1: Aerobic growth of E. coli
res_aero <- fba(mod)

# Finally, simulation 2: Anaerobic growth of E. coli
mod <- changeBounds(mod, react = "EX_o2_e", lb = 0) # before: -1000
res_anaero <- fba(mod)

# Report simulation results (i.e., growth and acetate production)
cat("[Aerobic growth]\n",
    " Growth rate:        ", res_aero@obj,"\n",
    " Acetate production: ", res_aero@fluxes[mod@react_id == "EX_ac_e"],"\n\n",
    "[Anaerobic growth]\n",
    " Growth rate:        ", res_anaero@obj,"\n",
    " Acetate prodcution: ", res_anaero@fluxes[mod@react_id == "EX_ac_e"],"\n",
    sep = "")
```

## Key differences to sybil

- cobrar is fully functional from reading SBML files until optimisation of linear programs, without the need of additional packages such as *sybilSBML* or *glpkAPI*.
- The GNU glpk library is a system requirement. There are no links/references to IBM's CPLEX in the *cobrar* package, which were one reason why *sybil* was discontinued on CRAN.
- [roxygen2](https://roxygen2.r-lib.org/) is used for documenting functions and classes.
- Feature trim: A range of functions in *sybil* are not part of *cobrar*. 
- In *cobrar*, R's garbage collector handles memory management, including memory associated to C++-objects and the pointers to these.
- Simplifications in class and function architecture
  - No more Class "SysBiolAlg" nor sub-classes. Different algorithms have their own function and detailed documentation of their return values
- In *sybil*, columns named "annotation" were actually concatenated CVTerms (https://synonym.caltech.edu/software/libsbml/5.20.0/cpp-api/class_c_v_term.html). To avoid confusion with other levels of annotation, the columns (e.g. in `react_attr` or `met_attr`) are named 'CVTerms' in *cobrar*.
- *cobrar* allows to assign SBOTerms to reactions, metabolites, genes.
- Performance (i.e., computation time) improvements in certain procedures:
  - identification of dead-end metabolites
  - reading/exporting SBML files,
  - pFBA algorithm (a.k.a MTF 'Minimization of Total Flux').
  - FVA; also now allows relaxed constraints on optimal growth (e.g. flux variability with 90-100% optimal growth)


## Notes

- *cobrar* exports SBML files level 3 version 2 with `fbc` version 3 and `groups` version 1. 
- Group assignments are only supported for reactions.
- currently only glpk is supported as solver. A plugin for IBM's ILOG CPLEX is planned.
- *Multiple objectives*. The SBML standard with its `fbc` extension allows to specify more than one objective (class `ListOfObjectives`). However, *cobrar* can only handle one current objective function per model, which is defined as an objective coefficient vector in slot `obj_coef` of an object of class `modelorg`. Note that when reading SBML models, *cobrar* will only use the first objective defined in the SBML document.

## References

1. <a id="R1"></a>Gelius-Dietrich, G., Desouki, A.A., Fritzemeier, C.J., Lercher, M.J. sybil – Efficient constraint-based modelling in R. BMC Syst Biol 7, 125 (2013). https://doi.org/10.1186/1752-0509-7-125
