---
title: "Additional constraints"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Additional constraints}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(cobrar)
```

## Constraint for a multi-carbon source growth environment

Standard FBA models already contain the linear constraints:

* Stationarity constraints: $\boldsymbol{S}\ \boldsymbol{v}=\boldsymbol{0}$
* Flux bounds on individual reactions: $\boldsymbol{lb} \le \boldsymbol{v} \le \boldsymbol{ub}$

In specific cases, it can be useful to include additional linear constraints to a model, e.g., to add additional thermodynamic constraints or interactions between reactions.

This tutorial aims to illustrate how additional constraints can be added to an existing model and to inspect their impact on flux distribution predictions.

First, we load a model of *Escherichia coli*'s core metabolism.

```{r}
fpath <- system.file("extdata", "e_coli_core.xml", package="cobrar")
mod <- readSBMLmod(fpath)
```

Second, we change the lower bounds of the exchange reactions for three potential carbon sources, namely Glucose, Fructose, and Fumarate. By doing so, we simulate a growth environment that provides three carbon sources simultaneously. The maximum uptake rate for each carbon source is set to $3\ mmol\  gDW^{-1}\ hr^{-1}$. Furthermore, we set the lower bound for the exchange reaction of oxygen (`EX_o2_e`) to $0\ mmol\ gDW^{-1}\ hr^{-1}$, to simulate anoxic conditions.

```{r}
mod <- changeBounds(mod, react = c("EX_glc__D_e","EX_fru_e", "EX_fum_e"),
                    lb = c(-3,-3,-3))
mod <- changeBounds(mod, react = "EX_o2_e", lb = 0)
```

Fumarate has 4 carbon atoms, while Glucose and Fructose have 6 carbon atoms each. Here, we want to add an additional constraint to the model that limits carbon source uptake to a total of 35 C-atom-mmol per gram dry weight and and per hour. This can be accomplished by:

```{r}
mod <- addConstraint(mod, react = c("EX_glc__D_e","EX_fru_e", "EX_fum_e"),
                     coeff = c(6,6,4), rtype = "L", lb = -35)
# print the user constraint
printConstraint(mod)
```
Next, we can perform flux balance analysis (`fba()`) with the additional constraint and inspect predicted fluxes of all exchange reactions.

```{r}
sol <- fba(mod)
getExchanges(mod, sol)
```


As wen can see, Fructose and Glucose are consumed, while the uptake of Fumarate is virtually zero. Also, there is a quite high production of protons (ID: `EX_h_e`). Next, we will add another constraint, that limits the release of protons depending on the growth rate. Specifically, we will limit the release to 5 mmol per newly formed gram dry weight biomass: $$v_{H^+} \le 5\ v_{Biomass}$$

Which is the same as: $$v_{H^+} - 5\ v_{Biomass} \le 0$$

This constraint can easily be added to the model:
```{r}
mod <- addConstraint(mod, react = c("EX_h_e","BIOMASS_Ecoli_core_w_GAM"),
                     coeff = c(1, -5), rtype = "U", ub = 0)
sol <- fba(mod)
getExchanges(mod, sol)
```
The results show that as an effect of the second constraint the proton production (`EX_h_e`) is reduced, and all three carbon sources are consumed, including Fumarate.


## Adding multiple constraints at once

In the above example, both constraints were added to the model successively, each with a separate `addConstraint()` call. You can achieve the same final outcome model by adding both constraints simultaneously:

```{r}
# Load model
fpath <- system.file("extdata", "e_coli_core.xml", package="cobrar")
mod <- readSBMLmod(fpath)

# Change bounds
mod <- changeBounds(mod, react = c("EX_glc__D_e","EX_fru_e", "EX_fum_e"),
                    lb = c(-3,-3,-3))
mod <- changeBounds(mod, react = "EX_o2_e", lb = 0)

# Adding constraints
mod <- addConstraint(mod,
                     react = list(
                       c("EX_glc__D_e","EX_fru_e", "EX_fum_e"),
                       c("EX_h_e","BIOMASS_Ecoli_core_w_GAM")
                     ),
                     coeff = list(
                       c(6,6,4),
                       c(1,-5)
                     ), rtype = c("L","U"), lb = c(-35,NA), ub = c(NA, 0))

# Perform FBA
sol <- fba(mod)
getExchanges(mod, sol)
```
