---
title: "Flux Variability Analysis (FVA)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Flux Variability Analysis (FVA)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Variabilibity in metabolite uptake/production of *E. coli*

This guide introduces Flux Variability Analysis (FVA) as applied in *cobrar* to the core metabolic network model of *Escherichia coli*. Here, FVA is applied to estimate possible flux ranges for metabolite exchanges (positive value = production, negative values = consumption) assuming a minimal media with D-glucose as sole carbon source and under aerobic or anaerobic growth conditions.

For the following example, two packages are required:

```{r setup}
library(cobrar)
library(ggplot2)
```

### Aerobic growth conditions

```{r, fig.dim = c(7, 4.8)}
fpath <- system.file("extdata", "e_coli_core.xml", package="cobrar")
mod <- readSBMLmod(fpath)

fvares <- fva(mod, react = mod@react_id[grepl("^EX_",mod@react_id)],
              opt.factor = seq(0.5, 1, length.out = 61))
fvares$react_name <- mod@react_name[match(fvares$react, mod@react_id)]
fvares$react_name <- gsub(" exchange$","",fvares$react_name)


ggplot(fvares,
       aes(x = react_name)) +
  geom_hline(yintercept = 0, linetype = 3) +
  scale_color_viridis_c() +
  geom_linerange(aes(ymin = min.flux, ymax = max.flux, col = growth.fraction),
                 linewidth = 6) +
  labs(x = "Metabolite exchange",
       y = "mmol / (gDW hr)",
       color = "Growth fraction") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1))
```

### Anaerobic growth conditions
```{r, fig.dim = c(7, 4.8)}
mod_anero <- changeBounds(mod, "EX_o2_e", lb = 0)

fvares <- fva(mod_anero, react = mod_anero@react_id[grepl("^EX_",mod_anero@react_id)],
              opt.factor = seq(0.5, 1, length.out = 61))
fvares$react_name <- mod_anero@react_name[match(fvares$react, mod_anero@react_id)]
fvares$react_name <- gsub(" exchange$","",fvares$react_name)


ggplot(fvares,
       aes(x = react_name)) +
  geom_hline(yintercept = 0, linetype = 3) +
  scale_color_viridis_c() +
  geom_linerange(aes(ymin = min.flux, ymax = max.flux, col = growth.fraction),
                 linewidth = 6) +
  labs(x = "Metabolite exchange",
       y = "mmol / (gDW hr)",
       color = "Growth fraction") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1))
```
